---
title: "Data Appendix to \"The role of homeownership in transit use in the Allston-Brighton neighborhood of Boston\""
author: "Victoria Tse"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools) 
library(stargazer)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F,
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")

#The following custom function simplifies the process of writing dfSummaries to html files
export_summary_table <- function(dfSummary_output){
  data_info <- attr(dfSummary_output, "data_info")
  ds_name <- data_info$Data.frame
  print(dfSummary_output,
      file = str_c("output/", ds_name, "_summary.html"),
      method = "browser",
      report.title = ds_name)
}
```

```{r set dfSummary css style, echo = F, include = F}
st_css()
```
# Appendix description

This Data Appendix documents the data used in "The role of homeownership in transit use in the Allston-Brighton neighborhood of Boston". It was prepared in a Rmarkdown document that contains both the documentation and the R code used to prepare the data used in the final estimation. It also includes descriptive statistics for both the original data and the final dataset, with a discussion of any issues of note.

The datasets used directly by the final analysis are saved in `processed-data/` at the end of this file.

# Raw data
*Each dataset you use will have its own documentation section. The next subsection in this document (Dataset description) is a template. You can copy this section and paste it into your document each time you need to add a section for a new dataset. Note that each line in the Dataset description section __must__ end with two spaces.* 
This section documents the datasets used in this analysis.

## Dataset description

**Citation:** 

Massachusetts Department of Transportation. (2012, June). Massachusetts Travel Survey.

**DOI:** 

N/A

**Date Downloaded:**  

3/7/20, from files provided by Susan Sayre

**Filename(s):** 

- HH.txt 
- PER.txt 
- PLACE.txt 


**Unit of observation:**  

Each unit is an individual person

**Dates covered:**  

6/14/10 - 10/24/11

### To obtain a copy

To obtain a copy, be a part of the Smith College institution and ask Professor Sayre! Also, agree to the necessary privacy measures to ensure the safety of the data. If not a part of Smith College as an institution, contact MassDOT.

### Importable version (if necessary)
N/A

**Filename(s):**  

- raw_data/filename.csv  
- HH.txt 
- PER.txt 
- PLACE.txt 
- VEH.txt


### Variable descriptions

- **RIBUS:** Use transit on a regular basis during the week  
	
	1 - Yes
	
	0- No 
 
- **HHVEH:** Number of household vehicles in working condition 

	Range 1-7 
	8 - 8 or more 

- **HISP:** Hipanic or Latino 
	
	1- Yes
	
	0- No 

- **RACE:** Ethnicity or race
	
	1 White alone 
	
	2 Black or African American alone 
	
	3 Indigenous/ Alaska Native alone
	
	4 Asian alone 
	
	5 Native Hawaiian/ Pacific Islander alone 
	
	6 Some other race alone 
	
	7 Two or more races 

- **OWN:** Homeownership 
	
	1 Own/Mortgaged 
	
	0 Renter

- **INCOME:** Household Income in 2009 
	
	1 Less than $15,000
	
	2 $15,00-$24,999
	
	3 $25,000-$34,999 
	
	4 $35,000- $49,999 
	
	5 $50,000-$74,999 
	
	6 $75,000-$99,999 
	
	7 $100,00- $149,999 
	
	8 $150,000 or more 

- **AGE:** Age in years  

- **LIC:** Valid Driver's License 
  
  1- Yes
  
  2- No

- **TRANS:** Have a transit pass? 

  1- Yes  
  
  0- No 

- **EMPLY:** Employed 

  1- Yes 
  
  0- No 
  
- **EDUCA:** Level of Education 

  1- Not a high school graduate, 12th or less (Includes very young children) 
  
  2-High school diploma or GED 
  
  3- Some college credit but no degree 
  
  4- Associate or technical school degree 
  
  5- Bachelor's or undergraduate degree 
  
  6- Graduate Degree (Includes professional degrees, MD, DDs, JD) 
  
- **ENROL:** Student 

  1- Yes 
  
  0- No 

- **BIKFA:** Bike facilities available at work/school? [ENROL = 1, WORK=1]  

  1- Yes 
  
  0- No 

  


### Data import code and summary

*Once you've described the variables, enter an R chunk by selecting Code -> Insert Chunk, or Ctrl+Alt+I, give it a name to describe the dataset you are importing. After importing, export a dataframe summary using the command.*

```{r import data 1}

hh_data <- read_csv(file.path("raw-data","mts-survey-data","HH.txt"), guess_max = 15000)
``` 

```{r import data 2}
per_data <- read_csv(file.path("raw-data","mts-survey-data","PER.txt"), guess_max = 15000)
```  

```{r import data 3}
place_data <- read_csv(file.path("raw-data","mts-survey-data","PLACE.txt"), guess_max = 15000)
``` 

`export_summary_table(dfSummary(dataset_name))`

*While it will make your resulting file long, you should not modify the chunk options to suppress printing of code and output. I would likely not include this in the documentation for an actual paper I was submitting, but including them here will let me read your code and the output message from R and may help identify data import concerns early in the process. Since these files will exist only electronically, their length is less of a concern. If you like to print out files to proofread and want me to help you shorten the printed versions, let me know. We can temporarily modify the chunk options for printing and restore them before you submit the assignment.*


```




# Data Processing and Combination
*This section should include a discussion of the processing and merging steps needed to create your basic data. The code to implement these steps should be included in chunks in this section. Once the final merged data has been created, you should use the dfSummary function again to summarize the data you will be using. You should also save a file containing all the objects you will use in your final analysis to the processed_data folder.* 

``` 
```{r Merge First 2 Data sets}
Dummy_1 <- merge(hh_data, per_data,  
                             by = "SAMPN")
``` 

```{r Merge Previous Data set with 3rd data set to get full data}
Preliminary_data <- merge(Dummy_1, place_data, 
                 by ="SAMPN")
``` 

```{r Isolate the Allston individuals}
  Final_data_half1 <- Preliminary_data[ which(Preliminary_data$ZIP=='02134'), ]
``` 

```{r Isolate the Brighton individuals}
  Final_data_half2 <- Preliminary_data[ which(Preliminary_data$ZIP=='02135'), ]
``` 

```{r Merge Allston and Brighton data sets}
Final_data <- rbind(Final_data_half1, Final_data_half2)

``` 


```{r Adjust variables}
Modified_final_data <- Final_data %>%
  
mutate(RIBUS = case_when(RIBUS== 1 ~ 1, 
                        RIBUS== 2 ~ 0),
      HISP = case_when(HISP== 1 ~ 1, 
                       HISP == 2 ~ 0),
      OWN = case_when(OWN== 1 ~ 1, 
                      OWN== 2 ~ 0),
      LIC = case_when(LIC== 1 ~ 1, 
                      LIC== 2 ~ 0),
      TRANS = case_when(TRANS== 1 ~ 1,
                        TRANS== 2~ 0),    
      EMPLY = case_when(EMPLY== 1 ~ 1,
                        EMPLY== 2 ~ 0),
      ENROL = case_when(ENROL == 1 ~ 1, 
                        ENROL== 2 ~ 0),
      BIKFA = case_when(BIKFA == 1 ~ 1, 
                        BIKFA == 2 ~ 0),
      RACE_factor = as.factor(case_when(RACE == 1 ~ "White", 
                            RACE== 2 ~ "Black", 
                            RACE== 3 ~ "Indigenous", 
                            RACE==  4 ~ "Asian", 
                            RACE== 5 ~ "Pacific Islander", 
                            RACE== 6 ~ "Other", 
                            RACE== 7 ~ "Multiracial")),
     INCOME_factor = as.factor(case_when(INCOME == 1 ~ "15", 
                              INCOME== 2 ~ "15-24", 
                              INCOME== 3 ~ "25-34", 
                              INCOME==  4 ~ "35-49", 
                              INCOME== 5 ~ "50-74", 
                              INCOME== 6 ~ "75-99", 
                              INCOME== 7 ~ "100-149", 
                              INCOME == 8 ~ "150" )),
  EDUCA_factor = as.factor(case_when(EDUCA== 1 ~ "K-12", 
                            EDUCA== 2 ~ "HS-GED", 
                            EDUCA== 3 ~ "Some_college", 
                            EDUCA==  4 ~ "Associates", 
                            EDUCA== 5 ~ "Bachelors", 
                            EDUCA== 6 ~ "Graduate" )))
```


```{r Create Sumary Statistics, echo= T, results= 'asis'}
Summary_statistics <- Modified_final_data %>% 
  select(RIBUS, OWN, HHVEH, LIC, TRANS, BIKFA, RACE, HISP, INCOME, EDUCA, ENROL, EMPLY, AGE) 

stargazer(Summary_statistics, type = "latex", 
          title = "Summary Statistics",
          header = F, 
          digits = 2) 

```


```{r Run regression 1 (just homeownership), results='asis'} 
Regression_1 <- lm(RIBUS ~ OWN, data = Modified_final_data)  

stargazer(Regression_1, type = "latex", 
          intercept.bottom = F,
          header = F)
``` 

```{r Run second regression (Transit availability), results= 'asis'}
Regression_2 <- lm(RIBUS ~ OWN + HHVEH + LIC + TRANS + BIKFA, data = Modified_final_data) 


stargazer(Regression_2, type = "latex", 
          intercept.bottom = F,
          header = F)

```  

```{r Run third regression (Race and ethnicity), results='asis'}
Regression_3 <- lm(RIBUS ~ OWN + factor(RACE) + HISP, data = Modified_final_data) 

stargazer(Regression_3, type = "latex", 
          intercept.bottom = F,
          header = F)

``` 

```{r Run fourth regression (Education), results= 'asis'}
Regression_4 <- lm(RIBUS ~  OWN + factor(EDUCA), data = Modified_final_data) 

stargazer(Regression_4, type = "latex", 
          intercept.bottom = F,
          header = F)

```

```{r Run fifth regresion (Income), results='asis'}
Regression_5 <- lm(RIBUS ~  OWN + factor(INCOME), data = Modified_final_data) 

stargazer(Regression_5, type = "latex", 
          intercept.bottom = F,
          header = F)
``` 

```{r Run sixth regression (Workers and Students), results='asis'} 
Regression_6 <- lm(RIBUS ~  OWN + EMPLY + ENROL, data = Modified_final_data) 

stargazer(Regression_6, type = "latex", 
          intercept.bottom = F,
          header = F)

``` 


```{r Run regression 7 (Age), results='asis'}
Regression_8 <- lm(RIBUS ~ OWN + AGE, data = Modified_final_data) 

stargazer(Regression_8, type = "latex", 
          intercept.bottom = F,
          header = F)

``` 

```{r Isolate homeowner data}
  Homeowner_data <- Modified_final_data[ which(Modified_final_data$OWN=='1'), ]
```

```{r Isolate Renter Data}
  Renter_data <- Modified_final_data[ which(Modified_final_data$OWN=='0'), ]
```

```{r Create Homeowner Summary Stats, echo= T, results= 'asis'}
Summary_statistics <- Homeowner_data %>% 
  select(RIBUS, OWN, HHVEH, LIC, TRANS, BIKFA, RACE, HISP, INCOME, EDUCA, ENROL, EMPLY, AGE) 

stargazer(Summary_statistics, type = "latex", 
          title = "Homeowner Summary Statistics",
          header = F, 
          digits = 2) 
```

```{r Create Renter Summary Stats, echo= T, results= 'asis'}
Summary_statistics <- Renter_data %>% 
  select(RIBUS, OWN, HHVEH, LIC, TRANS, BIKFA, RACE, HISP, INCOME, EDUCA, ENROL, EMPLY, AGE) 

stargazer(Summary_statistics, type = "latex", 
          title = "Renter Summary Statistics",
          header = F, 
          digits = 2) 
```


# Analysis Variables 
In depth descriptions of the variables and models can be found in the paper. The variables I adjusted were RIBUS, OWN, TRANS, LIC, BIKFA, HISP, EMPLY, ENROL. Each of them expressed "no" with a "2", which I adjusted to a "0". 
  



# Discussion of Data  
Some of the data may be slightly skewed by anomalous variables. For example, INCOME, despite having a 1-8 range, has a mean of 13.42 due to a 99 outcome for "Don't Know".








